---
title: "day_2"
author: "MS"
date: "07/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(magrittr)
library(stringr)


#data[[1]] <- c(3,4,3,1,2)
#data[[1]] <- c(3,5,1,5,3,2,1,3,4,2,5,1,3,3,2,5,1,3,1,5,5,1,1,1,2,4,1,4,5,2,1,2,4,3,1,2,3,4,3,4,4,5,1,1,1,1,5,5,3,4,4,4,5,3,4,1,4,3,3,2,1,1,3,3,3,2,1,3,5,2,3,4,2,5,4,5,4,4,2,2,3,3,3,3,5,4,2,3,1,2,1,1,2,2,5,1,1,4,1,5,3,2,1,4,1,5,1,4,5,2,1,1,1,4,5,4,2,4,5,4,2,4,4,1,1,2,2,1,1,2,3,3,2,5,2,1,1,2,1,1,1,3,2,3,1,5,4,5,3,3,2,1,1,1,3,5,1,1,4,4,5,4,3,3,3,3,2,4,5,2,1,1,1,4,2,4,2,2,5,5,5,4,1,1,5,1,5,2,1,3,3,2,5,2,1,2,4,3,3,1,5,4,1,1,1,4,2,5,5,4,4,3,4,3,1,5,5,2,5,4,2,3,4,1,1,4,4,3,4,1,3,4,1,1,4,3,2,2,5,3,1,4,4,4,1,3,4,3,1,5,3,3,5,5,4,4,1,2,4,2,2,3,1,1,4,5,3,1,1,1,1,3,5,4,1,1,2,1,1,2,1,2,3,1,1,3,2,2,5,5,1,5,5,1,4,4,3,5,4,4)


# data <- read.delim("../data/day_6_example_full.txt", header = F)
# 
# data %>% 
#   tidyr::separate(V1, sep = ",", into = as.character(c(1:26))) %>% 
#   write.csv("../data/day_6_bamboozled.csv")




```

WG100XM3

# Task 1

```{r}
data = list()
data[[1]]  = c(5)

n_days = 256

for(i in 2:(n_days+1)){
  data[[i]] <- replace(data[[i-1]], data[[i-1]]==0, 7)
  data[[i]] <- data[[i]] - 1
  
  n_new_fish <- sum(data[[i-1]] == 0)
  
  data[[i]] <- c(data[[i]], rep(8, times = n_new_fish))

  print(i)
}


```


# Task 2

```{r}
n_all_fish <- NULL
for(i in data){
  n_all_fish <- c(n_all_fish, length(i))
}
```

```{r}
plot_data <- data.frame(
  x = 0:(n_days), 
  y = n_all_fish
  ) 

plot_data %>% 
  ggplot2::ggplot(., aes(x = x, y = y)) + 
  geom_line() + 
  theme_light()

```

```{r}
value_128 <- plot_data[129, ]$y
value_128
```

```{r}
plot_data
```


9076200

```{r}
percent_growth <- function(initial, final, time_periods){
  ((final - initial)/ initial *100) / time_periods
}

exp_growth <- function(initial, time_periods, percent_growth){
  initial * (1 + percent_growth) * time_periods
}

exp_growth(initial = 300, 
           time_periods = 256, 
           percent_growth = 539689150680)
```


```{r}
percent_growth(initial = 1, 
               final = 348, 
               time_periods = 64)

percent_growth(initial = 1, 
               final = 90763, 
               time_periods = 128)
```


```{r}
plot_data

70907.81 /542.1875
```

```{r}
dat_test = NULL
for(i in 1:83){
  dat_test <-rbind(dat_test, 
                   data.frame(x = i, 
                              y_2 = generation_number(i)*n_original_generation))
}

dplyr::left_join(plot_data, dat_test, by = "x") %>% 
  dplyr::mutate(#by_five = y / 5, 
                off_by = y_2 - y, 
                #off_by_five = off_by / 5
                ) 

```


x = t 

exponential.model <- lm(log(Counts)~ Time)
summary(exponential.model)

timevalues <- seq(0, 30, 0.1)
Counts.exponential2 <- exp(predict(exponential.model,list(Time=timevalues)))

```{r}
exponential.model <- lm(log(y)~ x, data = plot_data[6:81, ])
#summary(exponential.model)

value <- 257
Counts.exponential2 <- exp(predict(exponential.model,list(x = value)))
Counts.exponential2 

for(i in 1:81){
  print(
    exp(predict(lm(log(y)~ x, data = plot_data[i:81, ]),
                list(x = value))) == 26984457539
  )
}

plot_data %>% 
  dplyr::filter(y %in% c(489, 163))


plot_data
```

 The term "exp(x)" is the same as writing ex or e^x or "e to the x" or "e to the power of x". 
 
 [ 9, 11]
 [17, 22]
 [32, 88]
-
 
[1, 5]

5 = ab^1

[81, 5934]

5934 = ab^81

-

5 = ab
a = 5/b

5934 = (5/b)*b^81

b = 1.0925208129979777893936542491309425357041785165297580510247510298 

a = 5 / b

a = 4.576572

```{r}
a = 4.576572
b = 1.0925208129979777893936542491309425357041785165297580510247510298 
x = 256
y = a*(b)^x

y
```


26984457539

```{r}
plot_data %>% 
  dplyr::mutate(change = y - lag(y)) %>% 
  ggplot2::ggplot(., aes(x = x, y = change)) + 
  #coord_cartesian(xlim = c(0, 50)) + 
  geom_line() + 
  theme_light()
```

```{r}
plot_data %>% 
  dplyr::mutate(change = y/lag(y)) %>% 
  ggplot2::ggplot(., aes(x = x, y = change)) + 
  #coord_cartesian(xlim = c(0, 50)) + 
  geom_line() + 
  theme_light()

plot_data %>% 
  dplyr::mutate(change = y/lag(y)) 
```

```{r}
plot_data %>% 
  dplyr::mutate(change = y/x) %>% 
  ggplot2::ggplot(., aes(x = x, y = change)) + 
  #coord_cartesian(xlim = c(0, 50)) + 
  geom_line() + 
  theme_light()

plot_data %>% 
  dplyr::mutate(change = y/x) 
```


```{r}
plot_data %>% 
  dplyr::mutate(change = y/y[1]) %>% 
  ggplot2::ggplot(., aes(x = x, y = change)) + 
  #coord_cartesian(xlim = c(0, 50)) + 
  geom_line() + 
  theme_light()

plot_data %>% 
  dplyr::mutate(change = y/y[1])
```
5, 14, 23, 32, 

1 - 5
32 - 88
64 - 1467
96 - 24939
128 - 423963
160 - 7207371
192 - 122525307
224 - 2082930219
256 - 35409813723

128 - 430418
192 - 126285000
256 = 

x / 126285000 = 293.4

```{r}
div <- c(1, 2, 4, 8, 16, 32, 64, 128, 256)

plot_data 
```

26984457539

```{r}
for(i in plot_data$x){
  print(paste(c(26984457539/i, i)))
}
```

```{r}
new_fish = c()
for(i in data){
  new_fish_i = sum(i[1:length(i)] == 8)
  new_fish = c(new_fish, new_fish_i)
}

plot_data %>% 
  dplyr::mutate(
    new_fish = new_fish
  ) %>% 
  ggplot2::ggplot(., aes(x = x, y = new_fish)) + 
  #coord_cartesian(xlim = c(0, 50)) + 
  geom_line() + 
  theme_light()
```


No: 

```{r}
data <- list()
data[[1]] <- c(3,4,3,1,2)
#data[[1]] <- c(3,5,1,5,3,2,1,3,4,2,5,1,3,3,2,5,1,3,1,5,5,1,1,1,2,4,1,4,5,2,1,2,4,3,1,2,3,4,3,4,4,5,1,1,1,1,5,5,3,4,4,4,5,3,4,1,4,3,3,2,1,1,3,3,3,2,1,3,5,2,3,4,2,5,4,5,4,4,2,2,3,3,3,3,5,4,2,3,1,2,1,1,2,2,5,1,1,4,1,5,3,2,1,4,1,5,1,4,5,2,1,1,1,4,5,4,2,4,5,4,2,4,4,1,1,2,2,1,1,2,3,3,2,5,2,1,1,2,1,1,1,3,2,3,1,5,4,5,3,3,2,1,1,1,3,5,1,1,4,4,5,4,3,3,3,3,2,4,5,2,1,1,1,4,2,4,2,2,5,5,5,4,1,1,5,1,5,2,1,3,3,2,5,2,1,2,4,3,3,1,5,4,1,1,1,4,2,5,5,4,4,3,4,3,1,5,5,2,5,4,2,3,4,1,1,4,4,3,4,1,3,4,1,1,4,3,2,2,5,3,1,4,4,4,1,3,4,3,1,5,3,3,5,5,4,4,1,2,4,2,2,3,1,1,4,5,3,1,1,1,1,3,5,4,1,1,2,1,1,2,1,2,3,1,1,3,2,2,5,5,1,5,5,1,4,4,3,5,4,4)
n_days = 132

for(i in 2:(n_days+1)){
  data[[i]] <- replace(data[[i-1]], data[[i-1]]==0, 7)
  data[[i]] <- data[[i]] - 1
  
  n_new_fish <- sum(data[[i-1]] == 0)
  
  data[[i]] <- c(data[[i]], rep(8, times = n_new_fish))
  
  # if(i >= 3){
  #   data[[i-2]] <- 0
  # }
  
  print(i)
}

data %>% length()
```



```{r}
generation_number <- function(day){2^(floor(day/9) +1)}
n_original_generation = 5
generation_number(83)*n_original_generation
```


```{r}
#seq(2, 90, 9)*5

data.frame(
  x = seq(1, 90, 7), 
  y = seq(1, 90, 7)*5
)




```

