---
title: "Untitled"
author: "MS"
date: "03/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(magrittr)
library(stringr)

data <- read.delim("../data/day_10_a.txt", col.names = "x", header = F)
```

# Task 1

- characters can never "face" each other unless they are of the same kind. 

{([(< {}*[*<>[]*}* >{[]{[(<()> - Expected ], but found } instead.
[[<*[*([])*)*<([[{}[[()]]]     - Expected ], but found ) instead.
[{[{*(*{}*]*{}}([{[{{{}}([]    - Expected ), but found ] instead.
[<(<(<(*<*{}*)*)><([]([]()     - Expected >, but found ) instead.
<{([(*[*[(<>()){}]*>*(<<{{     - Expected ], but found > instead.

Remove matching brackets one by one, see what's left and go from there

```{r}
for(i in 1:max(nchar(data$x))){

  data_i <- data %>% 
    dplyr::mutate(
      x = stringr::str_remove(x, "\\<\\>|\\(\\)|\\[\\]|\\{\\}")
    )
  
    # repeat until there is no change in the data
    if(sum(data_i == data) == nrow(data)) break
    
    data = data_i
  
}  

data %<>% 
  dplyr::mutate(
    illegal = str_extract(x ,"\\>|\\)|\\]|\\}"), 
    score = case_when(
      stringr::str_detect(illegal, "\\)") ~ 3, 
      stringr::str_detect(illegal, "\\]") ~ 57,
      stringr::str_detect(illegal, "\\}") ~ 1197,
      stringr::str_detect(illegal, "\\>") ~ 25137,
      TRUE ~ 0
    )
  )

sum(data$score, na.rm = T)
```

# Part 2 

Incomplete lines:

[({(<(())[]>[[{[]{<()<>> - Complete by adding }}]])})]
[(()[<>])]({[<{<<[]>>( - Complete by adding )}>]})
(((({<>}<{<{<>}{[]{[]{} - Complete by adding }}>}>))))
{<[[]]>}<{[{[{[]{()[[[] - Complete by adding ]]}}]}]}>
<{([{{}}[<[[[<>{}]]]>[]] - Complete by adding ])}>

[({([[{{ ... }}]])})] 

Select only the bits of data where score == 0, then complete the brackets but adding the missing brackets *in reversed order*. 

```{r}
reverse_and_score <- function(line){
  
  line %<>% 
    str_replace_all(., "\\(", "1") %>% 
    str_replace_all(., "\\[", "2") %>% 
    str_replace_all(., "\\{", "3") %>% 
    str_replace_all(., "\\<", "4")
  
  rev_line = rev(stringr::str_split(line , "")[[1]]) %>% 
    as.numeric()
  
  score = 0
  for(i in rev_line) score = score*5 + i
  
  return(score)
}

data %<>% 
  dplyr::filter(score == 0) %>% 
  dplyr::rowwise() %>% 
  dplyr::mutate(p2_score = reverse_and_score(x))

median(data$p2_score)
```
