---
title: "day_2"
author: "MS"
date: "07/12/2021"
output: html_document
---

Task 2 not completed. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(magrittr)
library(stringr)


#data[[1]] <- c(3,4,3,1,2)
data[[1]] <- c(3,5,1,5,3,2,1,3,4,2,5,1,3,3,2,5,1,3,1,5,5,1,1,1,2,4,1,4,5,2,1,2,4,3,1,2,3,4,3,4,4,5,1,1,1,1,5,5,3,4,4,4,5,3,4,1,4,3,3,2,1,1,3,3,3,2,1,3,5,2,3,4,2,5,4,5,4,4,2,2,3,3,3,3,5,4,2,3,1,2,1,1,2,2,5,1,1,4,1,5,3,2,1,4,1,5,1,4,5,2,1,1,1,4,5,4,2,4,5,4,2,4,4,1,1,2,2,1,1,2,3,3,2,5,2,1,1,2,1,1,1,3,2,3,1,5,4,5,3,3,2,1,1,1,3,5,1,1,4,4,5,4,3,3,3,3,2,4,5,2,1,1,1,4,2,4,2,2,5,5,5,4,1,1,5,1,5,2,1,3,3,2,5,2,1,2,4,3,3,1,5,4,1,1,1,4,2,5,5,4,4,3,4,3,1,5,5,2,5,4,2,3,4,1,1,4,4,3,4,1,3,4,1,1,4,3,2,2,5,3,1,4,4,4,1,3,4,3,1,5,3,3,5,5,4,4,1,2,4,2,2,3,1,1,4,5,3,1,1,1,1,3,5,4,1,1,2,1,1,2,1,2,3,1,1,3,2,2,5,5,1,5,5,1,4,4,3,5,4,4)



data <- read.delim("../data/day_6_a.txt", header = F)
# 
# data %>% 
#   tidyr::separate(V1, sep = ",", into = as.character(c(1:26))) %>% 
#   write.csv("../data/day_6_bamboozled.csv")
```

WG100XM3

# Task 1

```{r}
data = list()
data[[1]]  = c(5)

n_days = 80

for(i in 2:(n_days+1)){
  data[[i]] <- replace(data[[i-1]], data[[i-1]]==0, 7)
  data[[i]] <- data[[i]] - 1
  
  n_new_fish <- sum(data[[i-1]] == 0)
  
  data[[i]] <- c(data[[i]], rep(8, times = n_new_fish))

  print(i)
}

```

```{r}
lapply(data, write, "test.txt", append=TRUE, ncolumns=8000)

```


```{r}
sum(12:1)
```


# Task 2

(note for when you get there - you won't have to model this for every single fish. Just for groups of fish with different starting points, which can only be 5, 4, 3, 2, or 1)

If fish 1G1 (from the first generation) starts (day 0) at day 5, then a new fish is produced on
- day 6 + 7 + 7 + 7 + 7 + 7 + 7... 
  - on any given day thereafter, this fish would have produced `r floor(n_day/7) + 1` fish
  - so after 80 days, this fish produced `r floor(80/7) + 1` fish (n_g2 = 12)
  
    - fish baby 1G2 produces a new fish on:
      day 15 (day it was born 6 + 9) + 7 + 7 + 7 ... on any given day, this fish produced `r floor(n_day/7) - 1` fish. 
      After 80 days, this was n = 12
    - fish 2G2 produces a new fish on 
      day 6 + 7 + 9 = 22 + 7 + 7 + 7... `r floor(n_day) - 2` fish
      After 80, this was n = 9
    - fish 3G2 produces a new fish on day 6 + 7 + 7 + 9 = 29 + 7 + 7 + 7...  `r floor(n_day/7) - 3` fish
      After 80 days ... n = 8 
    - fish 4G2 6 + 3*7 + 9 = 36 + 7 + 7 + 7... 
      After 80 days `r floor(80/7) - 4`
      ...
      ...
    - fish 12G2 produces `r floor(80/7) - 12`
    - all G2 fish would produce (floor(80/7) * (floor(80/7) + 1)) - 12 - 11 - 10 - 9 - 8 - 7 - 6 - 5 - 4 - 3 - 2 - 1 ... 
      in other words: (n_g2 * floor(80/7)) - sum(12:1) = n_g3
      - First 1G3 fish is produced on day 15. it produces new fish on 15 + 9: 
                        day 24 + 7 + 7 + 7 + ... `r floor(80/7) - 2`
      - 2G3 on day 22 + 9 = 31 + 7 + 7 + 7 + ... `r floor(80/7) - 3` 
      - 3G3 on day 29 + 9 = 38 + 7 + 7 + 7 + ... `r floor(80/7) - 4`
      - 4G3 on day 36 + 9 = 45 + 7 + 7 + 7 + ... `r floor(80/7) - 5`
      - 5G3 on day 43 + 9 = 52 + 7 + 7 + 7 + ... `r floor(80/7) - 6`
      - 6G3 on day 50 + 9 = 59 + 7 + 7 + 7   ... `r floor(80/7) - 7`
      - 7G3 on day 57 + 9 = 66 + 7 + 7       ... `r floor(80/7) - 8`
      - 8G3 on day 64 + 9 = 73 + 7           ... `r floor(80/7) - 9`
      - 9G3 on day 71 + 9 = 80               ... `r floor(80/7) - 10`


From each generation, how many fish have time to reproduce? 

day_1G_repro: day on which the first fish from that generation produces a new fish.

n_repro = floor((80 - day_1G_repro)/7) + 1

how to figure out day_1G_repro? 

day_1G1_repro = 6
day_1G2_repro = 6 + 9
day_1G3_repro = 6 + 9 + 9

day_1G_repro = 6 + 9*(G-1)

... loop until a generation doesn't reproduce

```{r}
# day 0 of calendar is day 5 of the reproductive cycle
days = 80
cycle_count = floor(days/7)

n_g02 = cycle_count + 1
n_g03 = n_g02 * cycle_count - sum(n_g2: 1)
n_g04 = n_g03 * cycle_count - sum(n_g3: 2)
n_g05 = n_g04 * cycle_count - sum(n_g2: 3)
n_g06 = n_g05 * cycle_count - sum(n_g2: 4)
n_g07 = n_g06 * cycle_count - sum(n_g2: 5)
n_g08 = n_g07 * cycle_count - sum(n_g2: 6)
n_g09 = n_g08 * cycle_count - sum(n_g2: 7)
n_g10 = n_g09 * cycle_count - sum(n_g2: 8)
n_g11 = n_g10 * cycle_count - sum(n_g2: 9)
n_g12 = n_g11 * cycle_count - sum(n_g2:10)
n_g13 = n_g12 * cycle_count - sum(n_g2:11)
n_g14 = n_g13 * cycle_count - sum(n_g2:12)

total_fish = sum()
```

a) when does each generation produce the first fish?
b) how many fish from each generation reproduce 

```{r}
fish_repro_data = NULL
for(i in 1:20){
  
  G = i
  day_1G_repro = 6 + 9*(G-1)
  n_repro = floor((80 - day_1G_repro)/7) + 1
  
  fish_repro_data = rbind.data.frame(
    fish_repro_data,
    data.frame(
      G = G, day_1G_repro = day_1G_repro, n_repro = n_repro
    )
  )
  
}
```

```{r}
fish_repro_data
fish_repro_data %>% dplyr::filter(n_repro < 1)
```

In the span of 80 days 10th generation of fish is the last one that is produced. This generation does not produce any more fish.  

How many fish does each generation produce?

```{r}
# day 0 of calendar is day 5 of the reproductive cycle
days = 80
cycle_count = floor(days/7)

n_g00 = 1 # starter fish 
n_g01 = cycle_count
n_g02 = fish_repro_data[1, "n_repro"] * cycle_count - sum((fish_repro_data[1, "n_repro"]): 1)
n_g03 = fish_repro_data[2, "n_repro"] * cycle_count - sum((fish_repro_data[2, "n_repro"]): 2)
n_g04 = fish_repro_data[3, "n_repro"] * cycle_count - sum((fish_repro_data[3, "n_repro"]): 3)
n_g05 = fish_repro_data[4, "n_repro"] * cycle_count - sum((fish_repro_data[4, "n_repro"]): 4)
n_g06 = fish_repro_data[5, "n_repro"] * cycle_count - sum((fish_repro_data[5, "n_repro"]): 5)
n_g07 = fish_repro_data[6, "n_repro"] * cycle_count - sum((fish_repro_data[6, "n_repro"]): 6)
n_g08 = fish_repro_data[7, "n_repro"] * cycle_count - sum((fish_repro_data[7, "n_repro"]): 7)
#n_g09 = fish_repro_data[8, "n_repro"] * cycle_count - sum((fish_repro_data[8, "n_repro"]): 8)


sum(n_g00, n_g01, n_g02, n_g03, n_g04, n_g05, n_g06, n_g07, n_g08)*5

fish_repro_data
```




```{r}
n_all_fish <- NULL
for(i in data){
  n_all_fish <- c(n_all_fish, length(i))
}
```

```{r}
plot_data <- data.frame(
  x = 0:(n_days), 
  y = n_all_fish
  ) 

plot_data %>% 
  ggplot2::ggplot(., aes(x = x, y = y)) + 
  geom_line() + 
  theme_light()

```
